---
title: "BenGood Example"
author: "M. Lesperance, C. Tsao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BenGood Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

BenGood is a statistical package which is used to assess the goodness-of-fit of your
data to Benford's Law.
To learn more about the functions and statistics  in
this package, please refer to the function documentation and the corresponding 
paper.^[Lesperance M, Reed WJ, Stephens MA, Tsao C, Wilton B
(2016) Assessing conformance with Benford's Law: goodness-of-fit tests and
simultaneous confidence intervals. PLoS one; 11(3).]

This document demonstrates how to use the main functions in BenGood
through a simple example.

# Why use BenGood? 

Benford's Law is a probability distribution on first significant digits of 
numbers; for example, the first significant digit of 123 is 1. The law is 
particularly remarkable because many types of data are considered to be 
consistent with Benford's Law, and hence the law may be used as a tool to detect
errors in data collection or entry, or even data fraud and alterations. The law
has been used in many diverse areas, from diagnostic tests for mathematical
models in biology and genomics to fraud detection by government agencies.

BenGood is an
easy-to-use, statistically sound package that will help you assess whether or
not your dataset conforms to Benford's Law.

# The example: R code only

```{r, eval=FALSE}
library(BenGood)

# Create dataset
set.seed(123)  #optional, set random number generator seed for replicable results
x <- rnorm(100)

# Obtain first digit frequencies
freq <- firstdigitsfreq(x, 1)

CVMStats(freq) # Cramer-von Mises test statistics
LR.dec(freq) # Alternative: decreasing probabilities
LR.genben(freq) # Alternative: Generalized Benford
LR.rod(freq) # Alternative: Rodriguez's generalized Benford
LR.mat.mult(freq) # Alternative: general multinomial, i.e. non-negative probabilities

# 95% Simultaneous confidence intervals: for this example, Goodman and Sison
myCI <- SimultConf(freq, c('Goodman', 'Sis'), 0.05)

# Plot Goodman and Sison simultaneous CIs
plotCI(myCI$Lower, myCI$Upper)

```

\newpage

# The example: Output, with explanations

# Step 1: Setting up your data


```{r}
library(BenGood)

# Create dataset
set.seed(123) #optional, set random number generator seed for replicable results
x <- rnorm(100)

# Obtain first digit frequencies
freq <- firstdigitsfreq(x, 1)
# or equivalently, freqdigits(extractdigits(data))
```

Have your data in vector or matrix form in R. The first step to using BenGood is to 
extract the first significant digits from your dataset using `firstdigitsfreq`. 
These first digits will used in the rest of the BenGood functions.

# Step 2: Analysis

## Cramer-von Mises test statistics

The Cramer-von Mises tests included are tests of the null hypothesis that
the cell probabilities conform to Benford's Law versus 
the general alternative of multinomial probabilities.
P-values and their approximations, are included for three Cramer-von Mises tests:  W, U and A with corresponding
approximations, Wap, Uap and Aap. These test statistics are discrete analogues of the Carmer-von Mises,
Watson and Anderson-Darling statistics respectively.
The Pearson chi-squared goodness-of-fit is provided as X.

```{r}
CVMStats(freq)  # Cramer-von Mises test statistics
```


## Likelihood Ratio Tests

BenGood contains four likelihood ratio tests, testing the null hypothesis
of the cell probabilities following Benford's Law against varying alternative 
hypotheses, of: decreasing, Generalized Benford, Rodriguez's generalized
Benford, and general multinomial probabilities. 

All four tests are run below:

```{r}
LR.dec(freq) # Alternative: decreasing probabilities
LR.genben(freq) # Alternative: Generalized Benford
LR.rod(freq) # Alternative: Rodriguez's generalized Benford
LR.mat.mult(freq) # Alternative: general multinomial, i.e. non-negative probabilities
```

Important components of the output include `$SL`, the p-value generated, and
`$conv`, the convergence number indicates convergence of the numerical method
used to obtain the MLEs of the cell probabilities. A "0" indicates a successful 
convergence.

## Simultaneous Confidence Intervals

The simultaneous CI functions, `SimultConf` and `plotCI`, 
visualize departures of your data from Benford's Law and are highly recommended.

`SimultConf` generates Lower and Upper matrices containing the lower and upper
bounds of simultaneous $(1-\alpha)$% confidence intervals, where the probability
that the intervals simultaneously contain the corresponding Benford proportions is
approximately $(1-\alpha)$ if the data conforms to Benford's Law. 
The seven methods available are:

- Quesenberry and Hurst
- Goodman
- Bailey angular transformation
- Bailey square root transformation
- Fitzpatrick and Scott
- Sison and Glaz
- Univariate approximate Binomial confidence intervals

Here, 95% simultaneous confidence intervals are stored in `myCI` using
Goodman's formula, and Sison and Glaz's formula:
```{r}
# 95% Simultaneous confidence intervals: for this example, Goodman and Sison
# Lower and upper bound matrices
myCI <- SimultConf(freq, c('Goodman', 'Sis'), 0.05)
```

`plotCI` can then be used generate a plot of the intervals, given the lower and
upper bound matrices from `SimultConf`:

\newpage
```{r fig.align = 'center', fig.dim = c(5,5)}
# Plot Goodman and Sison simultaneous CIs
# Note: set argument legloc = TRUE if your machine supports locator 
# to choose the location of your legend
plotCI(myCI$Lower, myCI$Upper)
```


